---
title: "Illustration of L1-ball Prior"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require("l1ball")
```

## Sparse Regression

Generate the data  $n=200, p=1000$
```{r}
n = 200
p = 1000
X <- matrix(rnorm(n*p),n,p)
d =5
w0 <-    c(rep(0, p-d), rnorm(d)*0.1 +4)
y = X%*% w0 + rnorm(n,0,.5)
```

```{r}
fit_reg <- l1ball(y, X, steps = 1000,burnin = 2000)
```

Compare the posterior mean vs the oracle
```{r}
par(mfrow=c(1,2))  
plot(colMeans(fit_reg$trace_theta))
plot(w0,type='p',col='red')
```


Check the convergence of Markov chain

```{r}
par(mfrow=c(3,2))  
for(i in 1:3){
  plot((fit_reg$trace_theta[,p-d+i]))
  acf((fit_reg$trace_theta[,p-d+i]), main='')
}

```




##Change point detection

Generate the data
```{r}
p = 500
n = 500
d = 5
idx = floor(seq(1,n,length.out=(d+2))[2:(d+1)])
w0 <- rep(0, p)
w0[idx] =  -( rnorm(length(idx)) + 5)
X <- matrix(1, p, p)
X[upper.tri(X)] = 0
y = X%*%w0 + rnorm(p)*1
plot(y)
```

```{r}
fit_changepoint <- l1ball(y, X, steps = 1000,burnin = 2000)
```

Plot the Data vs Fitted (posterior mean)
```{r}

theta_mean <- colMeans(fit_changepoint$trace_theta)

plot(y)
lines(X%*%theta_mean,col="red", lwd=2)

```

Plot the Data vs a few random sample
```{r}

idx = sample(c(1:1000),3)
theta1 <- fit_changepoint$trace_theta[idx[1],]
theta2 <- fit_changepoint$trace_theta[idx[2],]
theta3 <- fit_changepoint$trace_theta[idx[3],]

plot(y, col = rgb(red = 0, green = 0, blue = 0, alpha = 0.8))
lines(X%*%theta1,col="red", lwd=2)
lines(X%*%theta2,col="blue", lwd=2)
lines(X%*%theta3,col="yellow", lwd=2)

```



## Linear trend filter

Generate the data


```{r}
n = 500
p = 500
d = 3

idx = floor(seq(1,n,length.out = (d+2))[2:(d+1)])

X= matrix(0,n,n)
for (i in c(1:n)){
  for (j in c(1:i)){
    X[i,j] = abs(j-i)+1
  }
}
X[,1]=1

# there are 4 line segments

idx= c(2,idx)

w0 = rep(0,p)
w0[idx] = (c(1,-2,1,-2))
psi = X%*%w0

y = psi + rnorm(n)*10
plot(y)
```

Fit the model (note: this requires longer run time)

```{r}
fit_linear_trend <- l1ball(y, X, steps = 1000,burnin = 5000)
```


Plot the Data vs Fitted (posterior mean)
```{r}

theta_mean <- colMeans(fit_linear_trend$trace_theta)

plot(y)
lines(X%*%theta_mean,col="red", lwd=2)

```




Plot the Data vs a few random sample
```{r}

idx = sample(c(1:1000),3)
theta1 <- fit_linear_trend$trace_theta[idx[1],]
theta2 <- fit_linear_trend$trace_theta[idx[2],]
theta3 <- fit_linear_trend$trace_theta[idx[3],]

plot(y, col = rgb(red = 0, green = 0, blue = 0, alpha = 0.8))
lines(X%*%theta1,col="red", lwd=2)
lines(X%*%theta2,col="blue", lwd=2)
lines(X%*%theta3,col="yellow", lwd=2)

```